---
title: "p8105_hw6_wc2788"
author: "Wen Cheng"
date: "11/29/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
library(modelr)
library(p8105.datasets)
```

## Problem 1
### Load and clean data
```{r}
birthweight <- read_csv("birthweight.csv") %>%
  janitor::clean_names() %>%
#convert numeric to factor
  mutate(babysex = factor(babysex,
                          levels = c(1, 2),
                          labels = c("male", "female")),
         frace = factor(frace, 
                        levels = c(1, 2, 3, 4, 8, 9), 
                        labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
         malform = factor(malform, 
                          levels = c(0, 1),
                          labels = c("absent", "present")),
         mrace = factor(mrace, 
                        levels = c(1, 2, 3, 4, 8), 
                        labels = c("White", "Black", "Asian", "Puerto Rican", "Other")))

 #check for missing data
sum(is.na(birthweight))
```
There are `r nrow(birthweight)` observations and `r ncol(birthweight)` variables in this dataset, and no missing value.

### Propose a regression model
bwt may depends on blength and mrace. We fit that initial model in the following code.
```{r}
fit = lm(bwt ~ blength + mrace, data = birthweight) 

#tidy output
summary(fit)
```

### Diagnostics
```{r}
birthweight %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = blength, y = resid)) + geom_point()
```

```{r}
birthweight %>% 
ggplot(aes(x = blength, y = bwt)) + geom_point()
```
longer baby length will have higher baby weight, so blength affects bwt.

```{r}
birthweight %>% 
  modelr::add_residuals(fit) %>% 
  ggplot(aes(x = mrace, y = resid)) + geom_boxplot()
```

```{r}
birthweight %>% 
ggplot(aes(x = mrace, y = bwt)) + geom_boxplot()
```
white mother will have higher weight baby compared with other race, black mother' baby will lower weight. So, mrace affects bwt.

### Hypothesis testing
```{r}
#null model are a subset of the terms in a more complex alternative model
fit_null = lm(bwt ~ blength + mrace, data = birthweight)
fit_alt = lm(bwt ~ blength + mrace + wtgain, data = birthweight)

anova(fit_null, fit_alt) %>% 
  broom::tidy()
```

### Make plot
```{r}
#model residuals against fitted values
birthweight %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```

### Compare model
```{r}
# One using length at birth and gestational age as predictors (main effects only)
model_1 <- lm(bwt ~ blength + gaweeks, data = birthweight)

# One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
model_2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight)
```

### Cross Validation
```{r}
#CV using modelr
cv_df = 
  crossv_mc(birthweight, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
#fit models and obtain RMSEs using mutate + map & map2
cv_df = 
  cv_df %>% 
  mutate(
   fit = map(train, ~lm(bwt ~ blength + mrace, data = .x)),
   model_1 = map(train, ~lm(bwt ~ gaweeks + blength, data = .x)),
   model_2 = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_fit = map2_dbl(fit, test, ~rmse(model = .x, data = .y)),
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y))
  )
```

### Make plot
```{r}
cv_df %>% 
  select(starts_with("rmse")) %>%              
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>%       
  ggplot(aes(x = model, 
             y = rmse, 
             color = model, 
             fill = model)) + 
  geom_violin(alpha = 0.5) +
  labs(x = "Model",
       y = "Root Mean Squared Error",
       title = "The RMSE Distribution For Each Model") 
```
The model_2 has the smallest RMSE, so it is the best fit in these three models. model_1 has the highest RMSE, the fit model has higher RMSE than model_2, so these two are not best fit.

## Problem 2

### Load data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

```{r}
#set a seed for reproducibility
set.seed(1234)

#draw 5000 bootstrap samples
bootstrap_samples =
  weather_df %>% 
  bootstrap(n = 5000) %>% 
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    log_beta = map(models, broom::tidy),
    r_squared =map(models, broom::glance)) %>% 
    select(.id, r_squared, log_beta) %>% 
    unnest(r_squared, log_beta)
```

### 95% CI for R Squared
```{r}
bootstrap_samples %>%
  summarize(
    lower_limit = quantile(r.squared, c(0.025)),
    upper_limit = quantile(r.squared, c(0.975))
  ) %>%
  knitr::kable()
```

### Make R Squared plot
```{r}
bootstrap_samples %>%
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(
    title = "Distribution of R Squared"
  )
```
The distribution of r squared peak is around 0.912 and it is a normal distribution. 

### 95% CI for log(beta0*beta1) 
```{r}
bootstrap_samples %>%
  select(.id, term, estimate) %>%
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    log_b0b1 = log(intercept*tmin)
  ) %>%
  summarize(
    lower_limit = quantile(log_b0b1, c(0.025)),
    upper_limit = quantile(log_b0b1, c(0.975))
  ) %>%
  knitr::kable()
```

### Make log(beta0*beta1) plot
```{r}
bootstrap_samples %>%
  select(.id, term, estimate) %>%
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    log_b0b1 = log(intercept*tmin)
  ) %>%
  ggplot(aes(x = log_b0b1)) + 
  geom_density() +
  labs(
    x = "log(beta0*beta1)",
    title = "Distribution of log(beta0*beta1)"
  )
```
The distribution of log(beta0*beta1) peak is around 2.017 and it is a normal distribution. 





